---
title: "Updating Lawler's Law"
output: github_document
---

```{r}
library(tidyverse)
library(RcppRoll)
```

```{r}
firsttox = read_csv('analysis/first-to-x-points.csv')
```
```{r}
head(firsttox)
```

### Shift of the scoring curve over time

```{r}
firsttox %>% 
  ggplot(aes(point, pct_correct, color = year)) +
  geom_point(alpha = 0.2) + 
  stat_smooth(method = 'lm', formula = y ~ x + I(x^2) + I(x^3))
```

### Then

Group 1997 to 2001, a five-year period about 20 years ago. Calculate some rolling averages for the percentage chance of a team winning for getting to that point total.

```{r}
firsttox %>% 
  filter(year <= 2001) %>% 
  group_by(point) %>% 
  summarise(games = sum(games), correct = sum(correct), pct_correct = correct / games) %>% 
  ggplot(aes(point, pct_correct)) +
  geom_line()
```

This trend is a little jumpy. The real trend should be monotonically increasing -- for each increment in the point value, the percentage chance that point total predicts the winner should not be lower than the point total before.

Calculate rolling totals for total games and correct games for a moving average.

```{r}
firsttox.97.01 = firsttox %>% 
  filter(year <= 2001) %>% 
  group_by(point) %>% 
  summarise(games = sum(games), correct = sum(correct)) %>% 
  mutate(
    period = '1997-2001',
    pct_correct = correct / games,
    games_roll = roll_sum(games, 5, align = "right", fill = NA),
    correct_roll = roll_sum(correct, 5, align = "right", fill = NA),
    pct_correct_roll = correct_roll / games_roll
  )
```

What's this trend look like?

```{r}
firsttox.97.01 %>% 
  ggplot(aes(point, pct_correct_roll)) +
  geom_line()
```

Is the percentage at each point increasing (or the same) based on the previous value?

```{r}
firsttox.97.01 %>% 
  select(point, pct = pct_correct_roll) %>% 
  mutate(prevpct = lag(pct), diff = pct - prevpct, diffpositive = diff >= 0) %>% 
  filter(!diffpositive)
```

It is. This is "monotonically increasing."

What was the percentage chance of scoring 100 first and winning during this period?

```{r}
firsttox.97.01 %>% filter(point == 100)
```

95.6% across 16,197 games.

### Now

Compare to the most recent five-year period: 2015-2019.

```{r}
firsttox.15.19 = firsttox %>% 
  filter(year >= 2015 & year <= 2019) %>% 
  group_by(point) %>% 
  summarise(games = sum(games), correct = sum(correct)) %>% 
  mutate(
    period = '2015-2019',
    pct_correct = correct / games,
    games_roll = roll_sum(games, 5, align = "right", fill = NA),
    correct_roll = roll_sum(correct, 5, align = "right", fill = NA),
    pct_correct_roll = correct_roll / games_roll
  )
```

Is this one monotonically increasing too?

```{r}
firsttox.15.19 %>% 
  ggplot(aes(point, pct_correct_roll)) +
  geom_line()
```

```{r}
firsttox.15.19 %>% 
  select(point, pct = pct_correct_roll) %>% 
  mutate(prevpct = lag(pct), diff = pct - prevpct, diffpositive = diff >= 0) %>% 
  filter(!diffpositive)
```

Yes.

What's the success rate of 100 here?

```{r}
firsttox.15.19 %>% filter(point == 100)
```

90% in 26,475 games. Makes sense: many more games and less success.

### The update

Compare the two monotonically increasing sequences. What was the percentage chance of 100 predicting the winner in the first five-year period? Take that value and find the first point value now that reaches that threshold.

```{r}
firsttox.15.19 %>% 
  arrange(point) %>% 
  filter(pct_correct_roll >= (firsttox.97.01 %>% filter(point == 100) %>% pull(pct_correct_roll))) %>% 
  slice(1)
```

The new magic number: 114.

How do the two curves compare?

```{r}
bind_rows(firsttox.97.01, firsttox.15.19) %>% 
  select(period, point, pct_correct_roll) %>% 
  ggplot(aes(point, pct_correct_roll, color = period)) +
  geom_line() +
  geom_hline(yintercept = 0.9559178) +
  geom_vline(xintercept = 114)
```
